# Chapter 2

## Page 12

### Normal Distribution

```{r}
#| include: false
library(tidyverse)
library(usethis) # use_github_file 
library(janitor) # clean_names() 
library(skimr) # skim

```

```{r}
library(tidyverse)

population_data <- data.frame(weights = rnorm(1000000, 80, 10))

weight_graph <- ggplot(data = population_data,
                   mapping = aes(x = weights)) +
  geom_density() +
  scale_x_continuous(name = "Distribution of Weights") +
  scale_y_continuous(name = "Density") +
  theme_classic()

print(weight_graph)
```

### Skewed Distribution

```{r}
library(tidyverse)

income_data <- data.frame(income = rf(1000000, df1 = 5, df2 = 2000))

income_graph <- ggplot(data = income_data,
                   mapping = aes(x = income)) +
  geom_density() +
  scale_x_continuous(name = "Distribution of Income", 
                     breaks = seq(0,  6, by = 2)) +
  scale_y_continuous(name = "Density",
                     breaks = seq(0, .8, by = .2)) +

  theme_classic()

print(income_graph)
```

## Page 27

### Standardized Scores

```{r}
sample1_oz <- c(40, 45, 50, 55, 60, 65, 70)
z_sample1_oz <- scale(sample1_oz, center = TRUE, scale = TRUE)


sample2_grams <- c(1100, 1150, 1200, 1400, 1700, 1725, 1775)
z_sample2_grams <- scale(sample2_grams, center = TRUE, scale = TRUE)

```

```{r}
#| echo: false

weight_data <- data.frame(sample1_oz, sample2_grams, z_sample1_oz, z_sample2_grams)

knitr::kable(weight_data)
```

## Page 30

### Correlation and Covariance

```{r}
sample1_oz <- c(40, 45, 50, 55, 60, 65, 70)
sample1_length = c(31, 33, 37, 38, 42, 45, 48)

cov_sample1 <- cov(sample1_oz, sample1_length)
print(cov_sample1)

correlation_sample1 <- cor(sample1_oz, sample1_length)
print(correlation_sample1)

```

But also note:

```{r}
# covariance is like correlation, but with the standard deviations included

correlation_sample1 * sd(sample1_oz) * sd(sample1_length)

# You get the same value as the covariance
print(cov_sample1)
```

## Page 34

### Nations 2018

#### Activate packages

```{r}
library(usethis) # use_github_file 
library(tidyverse) # read_csv 
library(janitor) # clean_names() 
library(skimr) # skim
```

#### Obtain data and save it to your computer

```{r}
#| include: false
use_github_file(repo_spec = "https://github.com/johnhoffmannVA/LinearRegression/blob/main/Nations2018.csv",
                save_as = "nations2018.csv")

nations2018 <- read_csv("nations2018.csv", show_col_types = FALSE) %>% clean_names()
```

```{r}
#| eval: false
use_github_file(repo_spec = "https://github.com/johnhoffmannVA/LinearRegression/blob/main/Nations2018.csv",
                save_as = "nations2018.csv")
```

#### Load data from your computer

```{r}
#| eval: false
nations2018 <- read_csv("nations2018.csv") %>% 
  clean_names()
```

#### Inspect data

```{r}
nations2018 %>% 
  glimpse()  
```

Or use

```{r}
#| eval: false
nations2018 %>% 
  view()  
```

```{r}
#| echo: false
knitr::kable(nations2018)
```

#### Descriptive with skim()

```{r}
library(skimr)

nations2018 %>% 
  skim()  
```

#### Descriptives with describe()

Alternatively, you could use the describe() command from the psych package as per the book. But, NEVER use library(psych) it will break the tidyverse. Instead use psyc:: before each psych package command.

```{r}
# psych package must be installed. But do not use library(psych)
# Notice how psych creates a mean for the nation column - which makes no sense
nations2018 %>% 
  psych::describe()  
```


### GSS 2018 

```{r}
library(usethis) # use_github_file 
library(tidyverse) # read_csv 
library(janitor) # clean_names() 
library(skimr) # skim
```

#### Obtain data and save it to your computer

```{r}
#| include: false
use_github_file(repo_spec = "https://github.com/johnhoffmannVA/LinearRegression/blob/main/GSS2018.csv",
                save_as = "gss2018.csv")

gss2018 <- read_csv("gss2018.csv", show_col_types = FALSE) %>% clean_names()
```

```{r}
#| eval: false
use_github_file(repo_spec = "https://github.com/johnhoffmannVA/LinearRegression/blob/main/GSS2018.csv",
                save_as = "gss2018.csv")
```

#### Load data from your computer

```{r}
#| eval: false
gss2018 <- read_csv("gss2018.csv") %>% 
  clean_names()
```

#### Inspect data

```{r}
gss2018 %>% 
  glimpse()  
```

#### skimr(): Describe focal variables

```{r}
library(skimr)

gss2018 %>% 
  select(pincome, female) %>%
  skim()  
```

#### psyc::describe(): Describe focal variables

```{r}
library(skimr)

gss2018 %>% 
  select(pincome, female) %>%
  psych::describe()  
```



#### Standard t.test

```{r}
t.test(gss2018$pincome ~ gss2018$female )
```

#### apaText t.test

```{r}
library(apaText)

# This code provides markdown text for Quarto documents

gss2018 %>%
  mutate(female = as.factor(female)) %>%
  apa.ind.t.test(female, pincome, var.equal = FALSE)

```

